---
title: "Final Project"
author: "Melvin Disla"
date: "7/1/2017"
output: html_document
---
## Analysis Project of North Carolina births
**Define four business objectives**
1.	Does the number of average visits to the doctor differ between married and unmarried moms?
2.	Does the number of average visits different between younger and mature moms?
3.	What is the impact that smoking vs. non-smoking has an effect on baby weight.
4.	Does smoking have a factor in premature births?

**Exploratory Data Analysis: Complete description of the Database, checks for outliers and invalid observations**
The data set that I will be using is the North Carolina births dataset. In 2004, the state of North Carolina released a large data set containing information on births recorded in this state. This data set is useful to researchers studying the relation between habits and practices of expectant mothers and the birth of their children. This data set has observations on 13 different variables; some are categorical and some numerical. The cases are the relations between the births of children in North Carolina and the habits of their expectant mothers. There are 1,000 cases. While looking over the dataset I came upon NA from female age, preemie, weeks, visits, marital, gained, habit, white mom columns. What I did was that I removed the NA from my functions so they won’t skew the answers. If I come up with any outliers in my data I would have to inspect it and determine the following: If it is obvious that the outlier is due to incorrectly entered or measured data, I would drop the outlier, If the outlier does not change the results but does affect assumptions, I will drop the outlier and make a note in my results, and lastly if the outlier creates a significant association, I will drop the outlier and not report any significance in my analysis. In my first objective the martial column is nominal categorical and the visits is numerical ordinal. Also there are some outliers in the visits column, since I am not sure of how the data was collected I will be reporting the histogram with and without the outliers and see which one correlates more. Just in the first objective, in the second objective used the same variable visit but to find the difference with the mature rate of mothers. The mature column is nominal categorical. When looking at the variables in objective three, baby weight is numerical continuous nominal. The variable habit is just ordinal categorical. With the variable weight is showing small outliers but is does not seem they were put in due to human error so I decided to leave them in for my analysis. The preemie column is ordinal and categorical.



**Data Visualization: Histograms, boxplots, median, mean, Quartiles, bar plots, scatterplots, pie charts, mosaic plots**
In my project I will be first getting a summary of each variable that I will be working with and that will give me the mean, mode, and the standard deviation. Next I will take a look at each data and see if it’s a categorical data I will make a bar plot and if it’s a numerical I will do a histogram, box plot and summarize about all the charts. I will check the skewness of the charts and check again via quantile plots summarize about them also. In the first objective the visit column it’s showing a right skewness for the histogram and the quantile plot. Then we have our boxplot showing the outliers, which I did two of them one with and without to see the difference to see if it was significant. The bars plot for martial shows that there is more not married mothers than married. Since in objective 2 I’ve used the same column of visit I will not report on it again. In objective three shows a left skewness with the two outliers in the histogram, quantile plot, and boxplot. The plots for the preemie column show that there are more full term babies than premature babies.

**Identify the appropriate set of hypotheses**
1. Ho: The number of average doctor visits equals the same between married and unmarried.
Ha: The number of average doctor visits doesn’t equal the same between married and unmarried.
2. Ho: The average of visits equals the same between younger and mature moms.
Ha: The average of visits does not equal the same between younger and mature moms.
3.Ho: The impact of smoking vs. non-smoking has no effect on baby weight
Ha: The impact of smoking vs. non-smoking has an effect on baby weight
4. Ho: Smoking has no factor in premature births
Ha: Smoking has a factor in premature births

**Identify the appropriate statistical conditions and if all of the Conditions have been met.** 
First I will check the three conditions required for applying the normal model: (1) independence, (2) at least 30 observations, and (3) the data are not strongly skewed. At this moment the conditions for normal model and inference are satisfied for all of my objectives. 
Since there satisfied I can move on the next step in doing the statistical test analysis. The ones that I will be using in this analysis will be the Z-score, T-test, and Chi-square. Since in the data set there isn’t a variable with multiple categorical answer I will not be able to do a ANOVA test but I can show similarity of an ANOVA test by doing a Chi-Square. Some of the objective may only have two of the statistical test analysis done to them. Then checking their confidence interval to see if they meet the 95% confidence interval.
Perform all statistical tests and ANOVA tables – calculation of p-values
In the calculation for Does the number of average visits to the doctor differ between married and unmarried moms? I chose to do a z-test which brought back a p-value = 0. Since objective 2 was looking for the difference like objective 1 I chose to do a z-test again and the p-value = 0.15 The calculation that I chose to do for objective three was a t-test which the p-value < 2.2e-16. When doing a chi-square test between the two variables of preemie and habit the p-value = 1.


**Include at least one set of confidence intervals**
The confidence interval for my first objective given the p-value=0, I reject the null hypothesis and conclude that the average number of visits to the doctor is different between married and non-married moms. So, given the p-value = 0.15 for objective 2, I fail to reject the null hypothesis and conclude that the average number of visits to the doctor is different between younger and mature moms. In a 95 percent confidence interval for objective three I reject the null hypothesis and accept the alternate hypothesis the impact of smoking vs. non-smoking has an effect on baby weight due to the p-value being extremely small. Since the p-value equal one shows that it’s non-significant and there is not basis for rejecting the null hypothesis that smoking has no factor in premature births.

**Correlation and Regression Analysis**
The correlation between visits and martial status is weak but going in the positive direction. The line regression equation for these two variables is Y^ = 9.0906 + 1.8647 and the R-squared is 0.05165. The correlation between visits and mature is negligible and going in the negative direction.  While looking at the line regression for objective 2 it’s Y^ = 12.6107 + -0.5828 and the R-squared:  0.001485, which shows no correlation. When it comes to the correlation between habit and weight they have a negligible correlation but that is going in the negative direction and when dealing with its line regression is Y^ = 7.14427 + -0.31554 and the R-squared: 0.003848. When looking at the p-value: 0.02779 of objective three line regression it enforces my decision to reject the null hypothesis and accept the alternate as I did for the t-test. When trying to find out if there is any correlation between habit and preemie variables comes up that their correlation is negligible and going in a negative direction. When taking a look at their line regression Y^= 1.127955 + -0.001478 with an R-squared: -0.001001, which also contributes to my no correlation.

## The Data
```{r}
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "nc.RData")
load("nc.RData")
head(nc)
tail(nc)
```

## Data Visualization
```{r}
summary(nc)
```

**Objective One**
```{r}
IQR(nc$visits, na.rm = TRUE)
hist(nc$visits)
barplot(nc$visits)
qqnorm(nc$visits)
qqline(nc$visits)
boxplot(nc$visits)
library(psych)
describe(nc$visits)
```
*removing Outliers*
```{r}
Visit <- table(log(nc$visits))
hist(Visit)
qqnorm(Visit)
boxplot(Visit)
```

```{r}
mar <- as.factor(nc$marital)
martial <- as.numeric(nc$marital)
married = subset(nc, marital=="married")
nonmarried = subset(nc, marital=="not married")
plot(nc$marital)
```

*Statistical test*
```{r}
inference(y=nc$visits, x=nc$marital, type="ht", est="mean", 
          null=0, method="theoretical", alternative="twosided")
```

*Correlation and Linear Regression*
```{r}
cor(nc$visits, martial, use = "complete.obs")
```

```{r}
regression <- lm(formula = nc$visits ~ martial, data = nc)
summary(regression)
plot(regression)
```



**Objective Two**
```{r}
IQR(nc$visits, na.rm = TRUE)
hist(nc$visits)
barplot(nc$visits)
qqnorm(nc$visits)
qqline(nc$visits)
boxplot(nc$visits)
library(psych)
describe(nc$visits)
```

*removing Outliers*
```{r}
Visit <- table(log(nc$visits))
hist(Visit)
qqnorm(Visit)
boxplot(Visit)
```

```{r}
matu <- as.factor(nc$mature)
mature <- as.numeric(nc$mature)
maturemom = subset(nc, marital=="mature mom")
youngermom = subset(nc, marital=="younger mom")
plot(nc$mature)
```

*Statistical Test*
```{r}
inference(y=nc$visits, x=nc$mature, type="ht", est="mean", 
          null=0, method="theoretical", alternative="twosided")
```

*Correlation and Linear Regression*
```{r}
cor(nc$visits, mature, use = "complete.obs")
```

```{r}
regression2 <- lm(formula = nc$visits ~ mature, data = nc)
summary(regression2)
plot(regression2)
```


**Objective Three**
```{r}
IQR(nc$weight)
hist(nc$weight)
barplot(nc$weight)
qqnorm(nc$weight)
qqline(nc$weight)
boxplot(nc$weight)
library(psych)
describe(nc$weight)
```

```{r}
hab <- as.factor(nc$habit)
habit <- as.numeric(nc$habit)
nonsmoker = subset(nc, habit=="non smoker")
smoker = subset(nc, habit=="smoker")
plot(nc$habit)
```

*Statistical Test*
```{r}
t.test(habit, nc$weight)
```

*Correlation and Linear Regression*
```{r}
cor(habit, nc$weight, use = "complete.obs")
```

```{r}
regression3 <- lm(formula = nc$weight ~ habit, data = nc)
summary(regression3)
plot(regression3)
```


**Objective 4**
```{r}
plot(nc$habit)
habit1 <- as.numeric(nc$habit)
```

```{r}
plot(nc$premie)
premie1 <- as.numeric(nc$premie)
```

```{r}
smoker = subset(nc, habit=="smoker")
pre = subset(nc, premie=="premie")
by(nc$premie, nc$habit, length)
hist(pre$smoker)
```

*Statistical Test*
```{r}
crosstab <- xtabs( ~ nc$habit + nc$premie , nc )
crosstab
chisq.test(crosstab)
```

*Correlation and Linear Regression*
```{r}
cor(habit1, premie1, use = "complete.obs")

```

```{r}
regression4 <- lm(formula = habit1 ~ premie1, data = nc)
summary(regression4)
plot(regression4)
```

